---
title: "Unicode Support"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Unicode Support}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)
```

```{r setup}
library(lofifonts)
```



## Unicode support

Each font supports a different number of glyphs.  The vector fonts include
just the ASCII character sets.

The font with the best unicode coverage is the bitmap font `unifont` which supports
over 50 thousand glyphs.


## Font overview

| Type   | Name            | Sizes                         | Unicode?              | # glyphs |
|--------|-----------------|-------------------------------|-----------------------|----------|
| Bitmap | Spleen          | 5x8, 6x12, 8x16, 12x24, 16x32 | Some                  | 450-1000 |
| Bitmap | Tamzen          | 5x9, 6x12, 7x13, 8x15, 8x16, 10x20 bold/regular | Some |  180    |
| Bitmap | Unifont         | 16x16                         | Yes. Plane 0, 1       | 113446   |
| Bitmap | Unscii          | 8x8, 8x16                     | Some                  | 3240     |
| Vector | gridfont        |                               | Lower case ASCII only | 68       |
| Vector | gridfont_smooth |                               | Lower case ASCII only | 68       |
| Vector | arcade          |                               | Upper case ASCII only | 66       |


## Rendering with `unifont`

```{r unicode, fig.height = 3, fig.width = 8}
txt <- "二項分布\xF0\x9F\x8E\xB2の英語表記は\n「Binomial distribution」である。"

bitmap_text_raster(txt, "unifont") |>
  plot()
```


## Default character

When a particular unicode glyph is not available, some fonts define
a default character which should be used instead.

If a default character is not defined for a font then unknown characters
will use `?`

## Codepoints supported by each font

Here is an example of looking for the `thumbs-down` and `thumbs-up` glyphs.

```{r}
codepoints <- c(0x1f44d, 0x1f44e)

unifont <- get_lofi_font('unifont')
unifont
# Unifont has both of these
codepoints %in% unifont$glyph_info$codepoint
bitmap_text_raster(intToUtf8(codepoints), 'unifont') |> plot()

spleen <- get_lofi_font('spleen-5x8')
spleen
# This spleen font has neither
codepoints %in% spleen$glyph_info$codepoint
bitmap_text_raster(intToUtf8(codepoints), 'spleen-5x8') |> plot()
```






